---
title: "pediatric_icu_beds_map"
author: "Stephanie Lamm"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(stringr)
library(readr)
library(sf)
library(tigris)
library(viridis)
library(cowplot)

# for leaflet
library(rgdal)
library("leaflet")
```

## Download Data 
Data obtained here: https://healthdata.gov/dataset/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/6xf2-c3ie 

Data is updated daily. Not all hospitals report all information every day. Check the "coverage" field to see the number of hospitals that reported data for that day. i.e. "inpatient_beds_coverage" is the number of hospitals reporting inpatient_beds in the state for that day.


```{r Download Hospital Capacity Data from CDC}

download.file("https://healthdata.gov/api/views/6xf2-c3ie/rows.csv", "COVID-19_Reported_Patient_Impact_and_Hospital_Capacity_by_State.csv")

hospital_capacity_data <-read.csv("COVID-19_Reported_Patient_Impact_and_Hospital_Capacity_by_State.csv")

```
# Pull in US States map from tigris
```{r Get U.S. States shapefile using Tigris}
us_states <- states(cb =TRUE, resolutio ="20m") %>%
	shift_geometry()

glimpse(us_states)

us_states %>%
	ggplot() + 
	geom_sf()

```

## Get the percent of pediatric ICU beds occupied in each state

```{r Pediatric ICU Beds }
pediatric_icu_data <- hospital_capacity_data  %>%  
  select(state, staffed_pediatric_icu_bed_occupancy, total_staffed_pediatric_icu_beds) %>%
  mutate(percent_pediatric_icu_beds_occupied =
           (staffed_pediatric_icu_bed_occupancy/total_staffed_pediatric_icu_beds)*100) %>%
  arrange(desc(percent_pediatric_icu_beds_occupied))
pediatric_icu_data
```

## Join the pediatric ICU data with the US States map from tigris

```{r join us_states with pediatric_icu data}
ped_beds_us <- left_join(us_states, pediatric_icu_data, by=c("STUSPS"= "state"))

```

## Map it
```{r}

pediatric_icu_occupancy_map <- ped_beds_us %>%
	ggplot() +
	geom_sf(aes(geometry=geometry, fill = percent_pediatric_icu_beds_occupied), color=NA) +
        geom_sf_label(aes(label = round(percent_pediatric_icu_beds_occupied,0))) +
				scale_fill_viridis(direction=-1) +
				theme_void() +
	      labs(title="Pediatric ICU Bed Occupancy", caption="Source:CDC")

```

## Save it 
```{r Save the map}
save_plot("pediatric_icu_occupancy_map.png", pediatric_icu_occupancy_map, base_height = NULL, base_width = 12)

```
